---
title: "CaseStudy2"
author: "Yucheol Shin"
date: "8/7/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=-1)

library(MASS)
library(tidyr)
library(dplyr)
library(caret)
library(class)
library(gridExtra)
library(readxl)
```

Data Prepartion
```{r}
caseStudy2 = read.csv("data/CaseStudy2-data.csv", header = TRUE)

caseStudy2$Attrition = as.factor(caseStudy2$Attrition)
caseStudy2$BusinessTravel = as.factor(caseStudy2$BusinessTravel)
caseStudy2$Department = as.factor(caseStudy2$Department)
caseStudy2$EducationField = as.factor(caseStudy2$EducationField)
caseStudy2$Gender = as.factor(caseStudy2$Gender)
caseStudy2$JobRole = as.factor(caseStudy2$JobRole)
caseStudy2$MaritalStatus = as.factor(caseStudy2$MaritalStatus)
caseStudy2$Over18 = as.factor(caseStudy2$Over18)
caseStudy2$OverTime = as.factor(caseStudy2$OverTime)
caseStudy2$EnvironmentSatisfaction  = as.factor(caseStudy2$EnvironmentSatisfaction )
caseStudy2$JobLevel = as.factor(caseStudy2$JobLevel)
caseStudy2$JobSatisfaction = as.factor(caseStudy2$JobSatisfaction)
caseStudy2$PerformanceRating = as.factor(caseStudy2$PerformanceRating)
caseStudy2$RelationshipSatisfaction = as.factor(caseStudy2$RelationshipSatisfaction)
caseStudy2$StockOptionLevel = as.factor(caseStudy2$StockOptionLevel)
caseStudy2$WorkLifeBalance = as.factor(caseStudy2$WorkLifeBalance)
caseStudy2$Attrition = ifelse(caseStudy2$Attrition=="No", 0, 1)
caseStudy2$Attrition = as.factor(caseStudy2$Attrition)
caseStudy2$OverTime = ifelse(caseStudy2$OverTime=="No", 0, 1)
caseStudy2$OverTime = as.factor(caseStudy2$OverTime)
str(caseStudy2)
```


Missing Data
```{r}
sapply(caseStudy2, function(x) sum(is.na(x)))
```
There is no missing data.


EDA Numeric Summary
```{r}
#Overtime is factor of 0 and 1. So when we make it to numeric, 0 becomes 1 and 1 becomes 2. Thus we minus 1.
caseStudy2 %>% group_by(Attrition) %>% 
  summarize(
    Mean_Income = mean(MonthlyIncome), 
    Mean_Years = median(YearsAtCompany), 
    Mean_OverTime = mean(as.numeric(OverTime) - 1), 
    Mean_Job_Satisfication = mean(as.numeric(JobSatisfaction)), 
    count = n())


caseStudy2 %>% group_by(JobSatisfaction) %>% 
  summarize(
    Mean_Income = mean(MonthlyIncome), 
    Mean_Years = median(YearsAtCompany), 
    Mean_OverTime = mean(as.numeric(OverTime) - 1), 
    Mean_Attrition = mean(as.numeric(Attrition) - 1), 
    count = n())

```


EDA Graph
```{r}

caseStudy2 %>% ggplot(aes(JobSatisfaction, JobRole)) + geom_count(aes(color = ..n.., size = ..n..)) + guides(color = 'legend') +
  labs(y="Job Role", 
       x="Job Satisfaction", 
       title="Job Role vs Satisfication")

caseStudy2 %>% ggplot(aes(JobSatisfaction, OverTime)) + geom_count(aes(color = ..n.., size = ..n..)) + guides(color = 'legend') +
  labs(y="Overtime", 
       x="Job Satisfaction", 
       title="Overtime vs Job Satisfication")
caseStudy2 %>% ggplot(aes(YearsAtCompany, color=JobSatisfaction)) + geom_freqpoly()+
  labs(y="Count", 
       x="Years at Company", 
       title="Job Satisfication over Years at Company")


caseStudy2 %>% ggplot(aes(Attrition, OverTime)) + geom_count(aes(color = ..n.., size = ..n..)) + guides(color = 'legend') +
  labs(y="Overtime", 
       x="Attrition", 
       title="Overtime vs Attrition")

caseStudy2 %>% ggplot(aes(Age, MonthlyIncome, color=Gender)) + geom_point() + geom_smooth(method="lm", se=F) +
  labs(y="Monthly Income", 
       x="Age", 
       title="Age vs Monthly Income by Gender")


caseStudy2 %>% ggplot(aes(MonthlyIncome, color=Attrition)) + geom_freqpoly()+
  labs(y="Count", 
       x="MonthlyIncome", 
       title="Monthly Income by Attrition")


```
1 WorkLifeBalance
2 NumCompaniesWorked
3 OverTimeYes
4 JobSatisfaction
Attrition           
MonthlyIncome
Gender               
YearsAtCompany       



Clean Constant feature
```{r}
# count number of unqiue values in column (1 is row, 2 is column)
apply(caseStudy2, 2, function(x) length(unique(x)))

cleanData <- subset(caseStudy2, select = -c(EmployeeCount, Over18, StandardHours))
```
We are dropping EmployeeCount, Over18, StandardHours features as there is only one unique value.




Multicollinearity for continous variables
We are using pearson correlation to find correlation between numeric data.
Data with 0.5 < perason correlation has strong correlation.
```{r}
numericColumns <- unlist(lapply(cleanData, is.numeric))  
numericData = cleanData[ , numericColumns]
correlation = cor(numericData, method = c("pearson"))

round_df <- function(x, digits) {
    numeric_columns <- sapply(x, mode) == 'numeric'
    x[numeric_columns] <-  round(x[numeric_columns], digits)
    x
}

correlation = round_df(correlation, 4)
```
Strong Correlation
TotalWorkingYears - Age
TotalWorkingYears - MonthlyIncome
TotalWorkingYears - YearsAtCompany
YearsInCurrentRole - YearsAtCompany
YearsInCurrentRole - YearsSinceLastPromotion
YearsInCurrentRole - YearsWithCurrManager
YearsAtCompany - YearsSinceLastPromotion
YearsAtCompany - YearsWithCurrManager
YearsWithCurrManager - YearsSinceLastPromotion


Chi-squared Test
Null = variables are independent
Alternative = there is a relationship
```{r}
tbl = table(cleanData$TotalWorkingYears, cleanData$Attrition) 
chisq.test(tbl) 

tbl = table(cleanData$YearsAtCompany, cleanData$Attrition) 
chisq.test(tbl) 

tbl = table(cleanData$YearsSinceLastPromotion, cleanData$Attrition) 
chisq.test(tbl) 

tbl = table(cleanData$YearsWithCurrManager, cleanData$Attrition) 
chisq.test(tbl) 

tbl = table(cleanData$YearsInCurrentRole, cleanData$Attrition) 
chisq.test(tbl) 

cleanData <- subset(cleanData, select = -c(ID, TotalWorkingYears, YearsSinceLastPromotion, YearsWithCurrManager,YearsInCurrentRole))
```
YearsSinceLastPromotion and YearsAtCompany has strong correlation. 
Since chi-squared test with YearsAtCompany reject the null, we can assume YearsAtCompany is more relate to Attrition than YearsSinceLastPromotion. Thus we choose YearsAtCompany for our model.
Since YearsAtCompany has lower p value and has strong correlation with other years variables, we choose YearsAtCompany for our variable.
Thus from multicorrlinearity, we got Age, MontlyIncome, YearsAtCompany.
Finally, we drop TotalWorkingYears, YearsSinceLastPromotion, YearsWithCurrManager, and YearsInCurrentRole. We also drop ID as that is not feature.


Stepwise Feature Selection
```{r}
model <- glm(Attrition ~., data = cleanData, family = binomial)
stepwise <- model %>% stepAIC(trace = FALSE)
summary(stepwise)
```


Cook's Distance for OutLiers
```{r}
cooksd <- cooks.distance(stepwise)

sample_size <- nrow(cleanData)
plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 4/sample_size, col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4/sample_size, names(cooksd),""), col="red")  # add label

influential <- as.numeric(names(cooksd)[(cooksd > (4/sample_size))])
cleanData <- cleanData[-influential, ]

model2 <- glm(Attrition ~., data = cleanData, family = binomial)
stepwise2 <- model2 %>% stepAIC(trace = FALSE)
summary(stepwise2)
```
From stepwise model we found
    Age + BusinessTravel + DailyRate + 
    Department + DistanceFromHome + Gender + HourlyRate + JobInvolvement + 
    JobLevel + JobSatisfaction + NumCompaniesWorked + OverTime + 
    PercentSalaryHike + RelationshipSatisfaction + StockOptionLevel + 
    TrainingTimesLastYear + WorkLifeBalance + YearsAtCompany
are the features we should use to build the model



Feature Importance
```{r}
importance <- varImp(stepwise2, scale=FALSE)
head(arrange(importance,desc(Overall)), n = 5)
head(importance)
```
Top 5 important features
1 WorkLifeBalance
2 NumCompaniesWorked
3 OverTimeYes
4 JobSatisfaction
5 StockOptionLevel


Prediction and Confusion Matrix
```{r}
set.seed(4)
splitPerc = .70
trainIndices = sample(1:dim(cleanData)[1],round(splitPerc * dim(cleanData)[1]))
train = cleanData[trainIndices,]
test = cleanData[-trainIndices,]


trainFit <- glm(Attrition ~., data = train, family = binomial)
trainModel <- trainFit %>% stepAIC(trace = FALSE)
pred <- predict(trainModel,test)
pred <- as.factor(as.numeric(pred>0.5))
confusionMatrix(pred, reference = test$Attrition)

```

Use original data to get the accuracy
```{r}
trainFit <- glm(formula = Attrition ~ Age + BusinessTravel + DailyRate + 
    Department + DistanceFromHome + Gender + HourlyRate + JobInvolvement + 
    JobLevel + JobSatisfaction + NumCompaniesWorked + OverTime + 
    PercentSalaryHike + RelationshipSatisfaction + StockOptionLevel + 
    TrainingTimesLastYear + WorkLifeBalance + YearsAtCompany, 
    family = binomial, data = train)
trainModel <- trainFit %>% stepAIC(trace = FALSE)

pred <- predict(trainModel,caseStudy2)
pred <- as.factor(as.numeric(pred>0.5))
confusionMatrix(pred, reference = caseStudy2$Attrition)
```

KNN
```{r}
set.seed(4)
splitPerc = .70
knnData <- caseStudy2
knnData[1:36] = sapply(knnData[,1:36], as.numeric)

trainIndices = sample(1:dim(knnData)[1],round(splitPerc * dim(knnData)[1]))
train = knnData[trainIndices,]
test = knnData[-trainIndices,]

knnModel = knn(train, test, train$Attrition, prob = TRUE, k = 5)
table(test$Attrition,knnModel)
CM = confusionMatrix(table(test$Attrition ,knnModel))
CM
```


KNN with original Data
```{r}

set.seed(4)
splitPerc = .70
knnData <- caseStudy2
knnData[1:36] = sapply(knnData[,1:36], as.numeric)

knnModel = knn(knnData, knnData, knnData$Attrition, prob = TRUE, k = 5)
table(knnData$Attrition, knnModel)
CM = confusionMatrix(table(knnData$Attrition ,knnModel))
CM

```


KNN with feature selection
```{r}

knnFeatureData <- subset(caseStudy2, select = c(Attrition, Age, BusinessTravel, DailyRate, Department, DistanceFromHome, Gender, HourlyRate, JobInvolvement, JobLevel, JobSatisfaction, NumCompaniesWorked, OverTime, PercentSalaryHike, RelationshipSatisfaction, StockOptionLevel, TrainingTimesLastYear, WorkLifeBalance, YearsAtCompany))
knnFeatureData[1:19] = sapply(knnFeatureData[,1:19], as.numeric)

set.seed(4)
splitPerc = .70

trainIndices = sample(1:dim(knnFeatureData)[1],round(splitPerc * dim(knnFeatureData)[1]))
train = knnFeatureData[trainIndices,]
test = knnFeatureData[-trainIndices,]

knnModel = knn(train, test, train$Attrition, prob = TRUE, k = 5)
table(test$Attrition,knnModel)
CM = confusionMatrix(table(test$Attrition ,knnModel))
CM

```


Salary Prediction
```{r}

data3 <- subset(caseStudy2, select = -c(ID, EmployeeCount, Over18, StandardHours, TotalWorkingYears, YearsSinceLastPromotion, YearsWithCurrManager,YearsInCurrentRole))
str(data3)
linearModel <- lm(MonthlyIncome ~., data=data3)  # build linear regression model on full data
stepwiseLinear <- linearModel %>% stepAIC(trace = FALSE)
cooksd <- cooks.distance(stepwiseLinear)

sample_size <- nrow(data3)
plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 4/sample_size, col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4/sample_size, names(cooksd),""), col="red")  # add label

influential <- as.numeric(names(cooksd)[(cooksd > (4/sample_size))])
data3 <- data3[-influential, ]

linearModel <- lm(MonthlyIncome ~., data=data3)  # build linear regression model on full data
stepwiseLinear <- linearModel %>% stepAIC(trace = FALSE)
summary(stepwiseLinear)


importanceIncome <- varImp(stepwiseLinear, scale=FALSE)
head(arrange(importanceIncome,desc(Overall)), n = 5)
head(importanceIncome)
```

Prediction and RMSE for Linear Regression
```{r}
set.seed(4)
splitPerc = .70
trainIndices = sample(1:dim(data3)[1],round(splitPerc * dim(data3)[1]))
train = data3[trainIndices,]
test = data3[-trainIndices,]


trainFit <- lm(MonthlyIncome ~., data=train)
trainModel <- trainFit %>% stepAIC(trace = FALSE)

#Predict monthly income
incomePred <- predict(trainModel, test)
head(incomePred)
#Model Summary
summary (trainModel)

RSS <- c(crossprod(trainModel$residuals))
#Mean squared error:
MSE <- RSS / length(trainModel$residuals)
#Root MSE:
RMSE <- sqrt(MSE)
RMSE
```


Predict Attrition with No Attrition Data
```{r}
attritionPredData = read.csv("data/CaseStudy2CompSet No Attrition.csv", header = TRUE)
attritionPredData$OverTime = ifelse(attritionPredData$OverTime=="No", 0, 1)
attritionPredData$OverTime = as.factor(attritionPredData$OverTime)
attritionPredData$EnvironmentSatisfaction  = as.factor(attritionPredData$EnvironmentSatisfaction )
attritionPredData$JobLevel = as.factor(attritionPredData$JobLevel)
attritionPredData$JobSatisfaction = as.factor(attritionPredData$JobSatisfaction)
attritionPredData$PerformanceRating = as.factor(attritionPredData$PerformanceRating)
attritionPredData$RelationshipSatisfaction = as.factor(attritionPredData$RelationshipSatisfaction)
attritionPredData$StockOptionLevel = as.factor(attritionPredData$StockOptionLevel)
attritionPredData$WorkLifeBalance = as.factor(attritionPredData$WorkLifeBalance)

trainFit <- glm(formula = Attrition ~ Age + BusinessTravel + DailyRate + 
    Department + DistanceFromHome + Gender + HourlyRate + JobInvolvement + 
    JobLevel + JobSatisfaction + NumCompaniesWorked + OverTime + 
    PercentSalaryHike + RelationshipSatisfaction + StockOptionLevel + 
    TrainingTimesLastYear + WorkLifeBalance + YearsAtCompany, 
    family = binomial, data = cleanData)
trainModel <- trainFit %>% stepAIC(trace = FALSE)

pred <- predict(trainModel,attritionPredData)
pred <- as.factor(as.numeric(pred>0.5))
attritionPredData$Attrition = pred
attritionPredResult <- subset(attritionPredData, select = c(ID, Attrition))
attritionPredResult$Attrition = ifelse(attritionPredResult$Attrition==0, "No", "Yes")
attritionPredResult$Attrition = as.factor(attritionPredResult$Attrition)
write.csv(x=attritionPredResult, file="Case2PredictionsShin Attrition.csv", row.names=FALSE,quote=FALSE)
```


Predict Salary with No MonthlyIncome Data
```{r}
salaryPredData <- read_excel( "data/CaseStudy2CompSet No Salary.xlsx")
salaryPredData$OverTime = ifelse(salaryPredData$OverTime=="No", 0, 1)
salaryPredData$OverTime = as.factor(salaryPredData$OverTime)
salaryPredData$Attrition = ifelse(salaryPredData$Attrition=="No", 0, 1)
salaryPredData$Attrition = as.factor(salaryPredData$Attrition)
salaryPredData$EnvironmentSatisfaction  = as.factor(salaryPredData$EnvironmentSatisfaction )
salaryPredData$JobLevel = as.factor(salaryPredData$JobLevel)
salaryPredData$JobSatisfaction = as.factor(salaryPredData$JobSatisfaction)
salaryPredData$PerformanceRating = as.factor(salaryPredData$PerformanceRating)
salaryPredData$RelationshipSatisfaction = as.factor(salaryPredData$RelationshipSatisfaction)
salaryPredData$StockOptionLevel = as.factor(salaryPredData$StockOptionLevel)
salaryPredData$WorkLifeBalance = as.factor(salaryPredData$WorkLifeBalance)

linearModel = lm(formula = MonthlyIncome ~ Age + Attrition + BusinessTravel + 
    JobLevel + JobRole + NumCompaniesWorked + OverTime + PercentSalaryHike + 
    PerformanceRating + YearsAtCompany, data = caseStudy2)
predSalary = predict(linearModel, newdata = salaryPredData)
salaryPredData$MonthlyIncome = predSalary
salaryPredResult <- subset(salaryPredData, select = c(ID, MonthlyIncome))
write.csv(x=salaryPredResult, file="Case2PredictionsShin Salary.csv", row.names=FALSE, quote=FALSE)
```